### 2018-05-26 related.tex
Revision from Prof. Xue

\section{Releated Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built upon
a high-performance GEMM implemention was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall strategy
to optimize memory access.

@@

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM are written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
directive-based programming approache.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
Both the auto-tuning and directive-based programming approaches
rely on the compiler to transform kernels in C to machine instructions.
Kernel performance can be improved with compiler techniques,
including SIMD vectorization~\cite{Larsen:2000,Zhou:2016,Zhou:2016b,
  Eichenberger2004,GCCSLP2007},
polyhedral optimization~\cite{Bondhugula2008A,Kong:2013}, 
and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall strategy of GEMM mainly concerns about
tiling factors, loop orders and parallelization.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first proposal to handle this problem.
\cite{gotogemm} gives a detailed discussion of different looping orders.
GEMM are usually parallelized only on layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows the users to specify a sophisticated configuration
that more than one loops of layer 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization can better fit complex architecture features
like multi-sockets and hyperthreading.

>>>>

\section{Related Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built 
on top of 
a high-performance GEMM implementation was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall workload-partitioning  strategy
to optimize memory access.

@@

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM is 
written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} use a 
directive-based programming approach.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
Both the auto-tuning and directive-based programming approaches
rely on the compiler to transform kernels in C to machine instructions.
Kernel performance can be improved with compiler techniques,
including SIMD vectorization~\cite{Larsen:2000,Zhou:2016,Zhou:2016b,
  Eichenberger2004,GCCSLP2007},
polyhedral optimization~\cite{Bondhugula2008A,Kong:2013}, 
and loop tiling~\cite{Lam1991,Spampinato:2014,Xue00}.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall workload-partitioning strategy used in
GEMM is mainly concerned with choosing 
tiling factors, loop orders and parallelization
techniques.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} 
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first work to address this problem.
In \cite{gotogemm}, a detailed discussion on
choosing
different loop orders is given.
GEMM are usually parallelized only at
layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows developers to specify a sophisticated configuration
so 
that more than one loops at layers 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization is well suited to complex architecture features
like multi-sockets and hyperthreading.

### 2018-05-26 scp.tex
fix the symbol of empty set
Revision from Prof. Xue

This section demonstrates the SCP method to handle inter-thread cache conflicts.
The basic idea is that inter-thread cache conflicts can be eliminated
if packed matrices used by different threads are fetched to different cache sets.
In Section~\ref{subsec:example}, we demonstrate SCP by an example.
Then a formal description is given in Section~\ref{subsec:formal}.

\subsection{Example}\label{subsec:example}

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e., $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g., $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g., $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tuning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of
various packed matrices.

\begin{table}
  \centering
  \caption{Sizes of packed matrices in DGEMM ($es = 8B$)}
  \label{tab:msizes}
  \begin{tabular}{cl|cl}
    \toprule
    matrix & size & matrix & size \\
    \midrule
    %% $A_2$, $A_3$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
    %% $A_4$ & $es M_r K_c = 8KB$   & $B_3$, $B_4$ & $es N_r K_c = 16KB$ \\
    $A_2$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
    $A_3$ & $es M_c K_c = 384KB$ & $B_3$ & $es N_r K_c = 16KB$ \\
    $A_4$ & $es M_r K_c = 8KB$   & $B_4$ & $es N_r K_c = 16KB$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Each $A_2$ is packed into a single continuous buffer.
Figure.~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
We see that every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure.~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets,
and inter-thread cache conflicts caused by $A_2$ are eliminated completely!

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partitioning styles of the shared cache,
referred to as way-partitioning and set-partitioning, respectively.
To guarantee the data layout in set-partitioning,
$A_2$ cannot be stored in a single continuous buffer any more.
Instead, $A_2$ is distributed to 12 continuous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partitioning of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continuous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collaboratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partitioning method,
or fall back to conventional way-partitioning for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
Memory layout of other packed matrices on lower layers are
determined by $A_2$ and $B_2$.
We assume that the share matrix $B_2$ is packed
in conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

Before diving into details, we make a few assumptions,
which are standard practice rarely violated,
on the architectures that SCP can deal with:
\begin{itemize}
\item The architecture has 2 or 3 levels of cache. %% i.e., $NL=2$ or $NL=3$
\item All caches are inclusive set-associative caches. %% i.e., $n_l \ge 2$ for $l \in [1, NL]$
\item Caches on the same level are homogeneous. %% i.e., $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
\item A cache's way-capacity is a multiple of the sum of way-capacity of its children,
i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
\end{itemize}

We use a memory descriptor $\mathcal{D}_t^M$ to specify 
the memory space occupied by packed matrix $M$ used by thread $t$.
The memory descriptor is essentially a subspace of
the whole address space, i.e., $\mathcal{D}_t^M \subset \mathcal{A}$.
%% The memory layout of a packed matrix is described by a memory descriptor,
%% which is essentially a subspace of the whole address space $\mathcal{A}$.
%% Each packed matrix $M$ used by thread $t$,
%% is associated with a memory descriptor $\mathcal{D}_t^M$
%% specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses $S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partitioning, the memory descriptor contains only one segment.
With set-partitioning used in previous section,
the memory descriptor contains a sequence of equally strided segments.

@@

SCP runs in three phases.
The first phase allocates memory buffers for all $A_2$ and $B_2$
instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
for storing private and shared matrices respectively.
The second phase computes memory descriptors for $A_2$ instances
by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
$\mathcal{D}_t^{A_2}$ , $t \in [0, NT)$.
Buffers of different threads are disjoint,
i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
The third phase computes memory descriptors for $B_2$.
Because is $B_2$ is shared among all threads,
$\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
The working flow of these three phases is listed
in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partitioning of shared caches.
At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierarchy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e., a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierarchy in a depth-first-search order.

The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
into $NT$ partitions, one for each thread.
First, the range of $\mathcal{D}_s$ is obtained
(line~\ref{line:lb}--\ref{line:ub}) and
the capacity of $\mathcal{D}_s$ is equally
divided among $NT$ threads (line~\ref{line:len}).
Then $\mathcal{D}_s$ is partitioned in a way-partitioning style
and each thread $t$ get buffer space for its part in $B_2$
(line~\ref{line:thread.for}--\ref{line:thread.forend}).

>>>>

This section introduces our SCP method for reducing
inter-thread cache conflicts.
The key insight is that inter-thread cache conflicts can be eliminated
if packed matrices used by different threads are fetched to different cache sets in a shared L2 cache.
In Section~\ref{subsec:example}, we illustrate 
SCP by an example.  In Section~\ref{subsec:formal},
we give the algorithms in our SCP method.

\subsection{An Example}\label{subsec:example}

Suppose we want to implement a double precision GEMM (DGEMM) on
the 4-core processor, where $es=8B$.
We now determine the tiling factors used based on
the constraints (\ref{eq:constraints.reg}) -- (\ref{eq:constraints.l3}).
Many solutions for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}), including
$(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. Auto
tuning returns
$M_r = 4$ and $N_r = 8$ as the best.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here, we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures and the program code,
it is reasonable to shrink $M_c$ slightly, with,
for example, $192$, $208$ and $224$ as good
candidates. Again, a turning process yields
$M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
and $N_c$ can be given a large value.
Here, we choose $N_c = 1024NT = 4096$.
With these tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} lists
the sizes of various packed matrices used.

\begin{table}
  \centering
  \caption{Sizes of packed matrices in DGEMM ($es = 8B$)}
  \label{tab:msizes}
  \begin{tabular}{cl|cl}
    \toprule
    Matrix & Size & Matrix & Size \\
    \midrule
    %% $A_2$, $A_3$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
    %% $A_4$ & $es M_r K_c = 8KB$   & $B_3$, $B_4$ & $es N_r K_c = 16KB$ \\
    $A_2$ & $es M_c K_c = 384KB$ & $B_2$ & $es N_c K_c = 8MB$ \\
    $A_3$ & $es M_c K_c = 384KB$ & $B_3$ & $es N_r K_c = 16KB$ \\
    $A_4$ & $es M_r K_c = 8KB$   & $B_4$ & $es N_r K_c = 16KB$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

Each $A_2$ is packed into a single continuous buffer.
Figure~\ref{fig:partition.conventional} shows how the
$A_2$ instances of 4 threads are distributed in the shared 16-way L2 cache.
Every cache set $s \in [0,2048)$ contains
data from all 4 $A_2$ instances,
potentially leading to inter-thread cache conflicts.
What if the $A_2$ matrices are distributed in the way shown in
Figure~\ref{fig:partition.segmented}?
$A_2$ of different threads live in strictly disjoint cache sets, enabling
the inter-thread cache conflicts caused by $A_2$ to be
eliminated completely!

Figures~\ref{fig:partition.conventional} and~\ref{fig:partition.segmented}
essentially represent two different partitioning styles for a shared cache,
referred to as \emph{way-partitioning} and \emph{set-partitioning}, respectively.
To obtain the data layout in set-partitioning,
$A_2$ can no longer be stored in a single continuous buffer.
Instead, $A_2$ will be distributed to 12 continuous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While avoiding inter-thread cache conflicts,
set-partitioning
introduces some complexity to GEMM implementations
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continuous ones.

What about $B_2$, the other packed matrix used in GEBP?
Unlike $A_2$, which is thread-private,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collaboratively by all threads.
There are two choices (with different tradeoffs), (1) privatize $B_2$ and apply 
set-partitioning and (2) 
fall back to the conventional way-partitioning 
for $B_2$.
$B_2$ can be made thread-private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but potentially increases the inter-thread conflicts caused by the shared $B_2$ matrix.
In this work,
these two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Algorithms}\label{subsec:formal}

Given the tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierarchy,
SCP will systematically determine the
the memory layout for the
packed matrices $A_2$ and $B_2$ used in GEMM.
The memory layout of the other packed matrices at
lower layers are determined by the memory layout of
$A_2$ and $B_2$.
We assume that the share matrix $B_2$ is packed
in the conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

we make a few standard assumptions, which usually
hold for the architecture considered by SCP:
\begin{itemize}
\item The architecture has 2 or 3 levels of cache; %% i.e. $NL=2$ or $NL=3$
\item All caches are inclusive set-associative caches; %% i.e. $n_l \ge 2$ for $l \in [1, NL]$
\item Caches at the same level are homogeneous; and %% i.e. $L_l^i = L_l^j$ for $i \ne j$, $l \in [1, NL]$.
\item A cache's way-capacity is a multiple of the sum of way-capacities of its children,
i.e., $(\frac{nt_l}{nt_{l-1}} \cdot wc_{l-1}) \mid wc_l$.
\end{itemize}

We use a memory descriptor $\mathcal{D}_t^M$ to specify 
the memory space occupied by a 
packed matrix $M$ used by thread $t$.
The memory descriptor is essentially a subspace of
the whole address space, i.e., $\mathcal{D}_t^M \subset \mathcal{A}$.
%% The memory layout of a packed matrix is described by a memory descriptor,
%% which is essentially a subspace of the whole address space $\mathcal{A}$.
%% Each packed matrix $M$ used by thread $t$,
%% is associated with a memory descriptor $\mathcal{D}_t^M$
%% specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses,
$S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partitioning, the memory descriptor contains only one segment.
With set-partitioning,
the memory descriptor contains a sequence of equally strided segments.

@@

SCP runs in three phases.
The first phase allocates memory buffers for all $A_2$ and $B_2$
instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
for storing private and shared matrices, respectively.
The second phase computes memory descriptors for $A_2$ instances
by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
$\mathcal{D}_t^{A_2}$ , where $t \in [0, NT)$.
Buffers of different threads are disjoint,
i.e., $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \emptyset$ if $i \ne j$.
The third phase computes memory descriptors for $B_2$.
Because is $B_2$ is shared among all threads,
$\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
The work flow of these three phases is listed
in Algorithms~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.

In Algorithm~\ref{alg:scp.phase1},
the per-thread sizes for memory buffers $\mathcal{D}_p$ and
$\mathcal{D}_s$ are computed
(lines~\ref{line:size.p} -- \ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Then $align$ is found
(lines~\ref{line:align.init} -- \ref{line:align.endfor}) 
and $size_p$ is enlarged to a multiple of $align$ (line~\ref{line:align}).
Specifically,
$align$ is initialized with the L1 cache line size $l_1$.
Then for each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
the earlier $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight on using $wc_l/nt_l$ is that
each way of the level-$l$ cache should be
equally partitioned among the $nt_l$ threads
in set-partitioning.
Finally, memory is allocated (lines~\ref{line:alloc.begin} -- \ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin} -- \ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} describes
the second phase of SCP.
The main component is a recursive procedure $subspace$
(lines~\ref{line:subspace.begin} -- \ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierarchy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
This procedure
has three input parameters, $l$ and $idx$ for
identifying the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all the children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e.,
a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$,
causing $subspace$ to return early
(lines~\ref{line:shortcut.begin} -- \ref{line:shortcut.end}).
Otherwise, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(lines~\ref{line:for.begin} -- \ref{line:for.end}).
For each child node $cidx$ at layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branch (lines~\ref{line:if} -- \ref{line:endif})
represents two distinct ways for computing
$\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierarchy in a depth-first-search order.

The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
into $NT$ partitions, one for each thread.
First, the range of $\mathcal{D}_s$ is obtained
(lines~\ref{line:lb} -- \ref{line:ub}) and
the capacity of $\mathcal{D}_s$ is equally
divided among the $NT$ threads (line~\ref{line:len}).
Then $\mathcal{D}_s$ is partitioned in a way-partitioning style
and each thread $t$ get buffer space for its part in $B_2$
(line~\ref{line:thread.for}--\ref{line:thread.forend}).
